\documentclass[answers]{exam}
%\documentclass{exam}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsthm,amssymb}
\usepackage{multicol}
\usepackage{mathpazo}

\title{CHEM 1100 Practice Exam}
\author{Richard Robinson}

\begin{document}
\begin{center}
\fbox{\fbox{\parbox{5.5in}{\centering
Comprehensive practice exam for CHEM 1100. All questions based on previous assessment and thus contains $>90\%$ of the actual exam.}}}
\end{center}
\header{CHEM 1100 Practice Exam}{}{Richard Robinson}
\footer{}{Page \thepage\ of \numpages}{}
\setlength\columnsep{20pt}

%---------------------------------------------------------------
% MULTIPLE CHOICE
%---------------------------------------------------------------

\section{Multiple Choice (1)}
\begin{multicols}{2}
\begin{questions}

\question Which of the following is an example of a homogenous equilibrium?

\begin{choices}
 \choice $\mathrm{MgCO_3(s) \leftrightharpoons MgO(s) + CO_2(g)}$
 \choice $\mathrm{NaCl(s) \leftrightharpoons Na^+(aq) + Cl^-(aq)}$
 \CorrectChoice $\mathrm{3H_2(g) + N_2(g) \leftrightharpoons 2NH_3(g)}$
 \choice $\mathrm{C(s) + CO_2(g) \leftrightharpoons 2CO(G)}$
 \choice None of the above.
\end{choices}

\question A system in chemical equilibrium is \emph{not} characterized by one of the following:

\begin{choices}
  \choice Dynamic interconversion between reactants and products
  \choice No macroscopic changes
  \CorrectChoice Unaffected by changes in temperature
  \choice Unaffected by addition of catalyst
  \choice None of the above
\end{choices}

\question Which is \emph{false} about the first law of thermodynamics?

\begin{choices}
  \choice All energy change in a chemical reaction is in the form of heat
  \choice The enthalpy of the universe is zero
  \choice $\Delta E = \Delta E_\text{sys} + \Delta E_{\text{sur}}$
  \choice Energy cannot be created nor destroyed
  \choice The combined amount of matter in the universe is constant
\end{choices}

\question Which is \emph{false} about the second law of thermodynamics?

\begin{choices}
  \choice In any spontaneous process, entropy of the universe increases
  \choice In any spontaneous process, entropy of the system increases
  \choice The entropy of the surroundings can increase or decrease
  \choice $\Delta S_u = \Delta S + \Delta S_s$
  \choice The entropy of the universe is positive for a spontaneous process
\end{choices}

\question If $\mathrm{Ba(NO_3)_2}$ is added to $\mathrm{BaSO_4}$, the solubility of the latter:

\begin{choices}
  \choice is unaffected
  \choice is unpredictable
  \choice decreases
  \choice increases
\end{choices}

\question Will Cu have a higher conductivity at 200$^\circ$ than 100$^\circ$?

\begin{oneparchoices}
  \choice Yes
  \CorrectChoice No
\end{oneparchoices}

\question Will Si have a higher conductivity at 25 K than 75 K?

\begin{oneparchoices}
  \choice Yes
  \CorrectChoice No
\end{oneparchoices}

\question In an $n$-type semiconductor, does it take less energy for an electron to jump from the donor level or valence band to the conduction band?

\begin{oneparchoices}
  \CorrectChoice Donor
  \choice Valence
\end{oneparchoices}

\question Will an undoped semiconductor have a higher conductivity than an $n$-type semiconductor?

\begin{oneparchoices}
  \choice Yes
  \CorrectChoice No
\end{oneparchoices}


\end{questions}
\end{multicols}

%---------------------------------------------------------------
% SHORT ANSWERS
%---------------------------------------------------------------

\newpage
\section{Short Answer (2)}
\begin{multicols}{2}
\begin{questions}

\question Calculate $\Delta S$ for the reaction $\mathrm{2NO_2 \to 2N_2 + O_2}$. Note $\Delta S = \{ 240, 191.5, 205 \}$, respectively.
\answerline[108]

\question What is the total number of lone pairs in $\mathrm{NCl_3}$?
\answerline[10]

\question In manufacturing steel, carbon is likely to be a (?) impurity because it is (?) than iron.
\answerline[int., smaller]

\question In the reaction $\mathrm{A(g) + 3 B(\ell) \to 3 C(g) + 7 D(g)}$, what are the exponents in the denominator of the equilibrium expression?
\answerline[1; 0]

\question Given the heat of formation values $\{-103.8, 0, -393.5, -285.8 \}$, calculate the heat of reaction for $\mathrm{A(g)}$ $\mathrm{ + B(g) \to C(g) + D(\ell)}$.
\answerline[$\mathbf{-2.22 \cdot 10^3}$]

\question What is the molar solubility of $\mathrm{CaF_2}$ if $K_{sp} = 3.9 \cdot 10^{-11}$?
\answerline[$\mathbf{2.14 \cdot 10^{-4}}$]

\question What is the pH of a 0.15 M NaOH solution?
\answerline[13.18]

\question Which type of solid is most densely packed?
\answerline[fcc]

\question What element (Ga, Si, Al, Ar) would be added to Ge to produce an \textit{n}-type conductor?
\answerline[Ar]

\question The volume of a gas is 650 mL at STP. What volume will it occupy at freezing point and 950 torr?
\answerline[520]

\question For the reaction $\mathrm{H_2O(g) + CH_4(g) \leftrightharpoons CO(g)}$ $+ 3H_2(g)$, if $K_c = 3.8 \cdot 10^{-3}$ at 1000 K, what is $K_p$?
\answerline[26]

\question For the reaction $\mathrm{PCl_5(g) \leftrightharpoons PCl_3(g) + Cl_2(g)}$, the constant $K = 7.7 \cdot 10^{-3}$ at STP. Calculate $\Delta G^\ominus$.
\answerline[12,000 J/mol]

\question Predict whether the entropy change is positive or negative for the following processes:
\begin{parts}
  \part $\mathrm{CH_3OH(l) + 3/2 O_2(g) \to CO_2(g) + 2H_2O}$
  \answerline[Positive]
  \part $\mathrm{Na(s) + 1/2 F_2(g) \to NaF(s)}$
  \answerline[Negative]
  \part Compressing gas from 0.5 to 1.5 atm.
  \answerline[Negative]
  \part Heating copper by 20 K
  \answerline[Positive]
\end{parts}

\question If $\Delta G_f^\ominus$ for $CO, H_2, CH_4, H_2O$ are $-137.2, 0,$ $-50.75, -228.6$, calculate $\Delta G_{rxn}^\ominus$ for $\mathrm{CH_4 + H_2O}$ $\mathrm{\to CO + 3H_2}$.
\answerline[142.2 kJ]

\question For a sample which is 15 mol \% polymer with a molecular weight $M = 500,000$ g/mol and 85 mol \% with $M = 150,000$, what is the number average molecular weight?
\answerline[202,500]

\question What is the DP for $C_8H_8$ with a molecular weight of 100,000 g/mol?
\answerline[960]

\question Give an example of a ceramic.
\answerline[CSi]

\question For the reaction $\mathrm{2H_2SO_4 + Cu \to CuSO_4 +}$ $\mathrm{2H_2O + SO_2}$, how much $SO_2$ is produced from 14.2g Cu and 18g $H_2SO_4$?
\answerline[5.88g]

\end{questions}
\end{multicols}

%---------------------------------------------------------------
% LONG ANSWERS
%---------------------------------------------------------------

\newpage
\section{Long Answer (6)}
\begin{questions}

\question Use the Born Haber cycle to determine the lattice energy of KF (s) from the following data:
\begin{align*}
  \Delta H_f^\ominus &= - 567.3 & \Delta H_{sub} [K(s)] &= 89.24 & \Delta H_{dis} [F_2(g)] &= 159 \\
  IE[K(g)] &= 418.9 & EA[F(g)] &= -328
\end{align*}
%
\begin{solutionbox}{3.2in}
  The Born Haber Cycle is given by \begin{equation*}
    \Delta H_f^\ominus = \sum \Delta H^\ominus
  \end{equation*}
  in which the enthalpies are given by
  \begin{align*}
    & \text{Formation} & K(s) + \textstyle\frac{1}{2} F_2(g) &\to KF(s) & \Delta H_f^\ominus &= -567.3 \\
    & \text{Sublimation} & K(s) &\to K(g) & \Delta H_s &= 89.24 \\
    & \text{Ionization} & K(g) &\to K^+(g) + e^- & \Delta H_i &= 418.9 \\
    & \text{Dissociation} & \textstyle\frac{1}{2} F_2 (g) &\to F(g) & \Delta H_d &= 0.5 \cdot 159 \\
    & \text{Affinity} & F(g) + e^- &\to F^- (g) & \Delta H_e &= -328
  \end{align*}
  Therefore, the cycle is \begin{equation*}
    -567.3 = 89.24 + 418.9 + 0.5 \cdot 159 - 328 - \Delta H_l \implies \Delta H_l = 827
  \end{equation*}
  The lattice energy is thus 827 kJ/mol.
\end{solutionbox}

\bigskip
\question For the reaction $H_2(g) + I_2(g) \leftrightharpoons 2HI(g)$, the constant $K = 57$ at 700K. If 1 mol $H_2$ reacts with 1 mol $I_2$ in a 10L vessel at 700K, what is the molar composition at equilibrium?
%
\begin{solutionbox}{3.05in}
  The initial concentrations are given by $[H_2] = [I_2] = n/V = 0.1 \text{ M}$. An ICE table is next constructed to determine the concentrations at equilibrium:
  \begin{center}
  \begin{tabular}{llll}
    R & $H_2$ & $I_2$ & $2HI$ \\[1mm]
    I & 0.100 & 0.100 & 0 \\[1mm]
    C & $-x$ & $-x$ & $+2x$ \\[1mm]
    E & $0.100 - x$ & $0.100 - x$ & $2x$
  \end{tabular}
  \end{center}
  The change in concentration is therefore given by \begin{equation*}
    K_c = 57 = \frac{(2x)^2}{(0.1-x)(0.1-x)} \implies x = \begin{cases}
      0.0791 \text{ M} \\ 0.136 \text{ M}
  \end{cases}
  \end{equation*}
    Thus, the valid solution is $x = 0.0791$ so the concentrations at equilibrium are \begin{equation*}
      [H_2]_{eq} = [I_2]_{eq} = 0.1 - x = 0.0209 \text{ M} \quad\text{and}\quad [HI]_{eq} = 2x = 0.1582 \text{ M}
    \end{equation*}
\end{solutionbox}

\bigskip
\question A mixture of 1.57 mol $N_2$, 1.92 mol $H_2$, and 8.13 mol $NH_3$ is mixed in a 20L vessel at 500K. At this temperature, $K_c = 1.7 \cdot 10^2$ for $N_2 + 3H_2 \leftrightharpoons 2NH_3$. Is such mixture at equilibrium? If not, what is the direction of the net reaction?
%
\begin{solutionbox}{1.95in}
  The initial concentrations are given by \begin{equation*}
    [N_2] = 0.0785 \qquad [H_2] = 0.0960 \qquad [NH_3] = 0.406
  \end{equation*}
  The reaction quotient is thus \begin{equation*}
    Q = \frac{[NH_3]^2}{[N_2][H_2]^3} = \frac{(0.406)^3}{(0.0785)(0.0960)^3} = 2.37 \cdot 10^3
  \end{equation*}
  Therefore, $Q > K$ so the mixture is not at equilibrium and the net reaction will proceed leftwards, decreasing the $NH_3$ concentration.
\end{solutionbox}

\bigskip
\question Define a formation reaction, and explain why the each equation is not correct:
\begin{parts}
  \part The formation of $\mathrm{4Al(s) + 3O_2(g) \to 2 Al_2 O_3 (s)}$
  \part The formation of $\mathrm{N_2(g) + 3/2 H_2(g) \to NH_3(g)}$
  \part The formation of $\mathrm{2Na(s) + O(g) \to Na_2O(s)}$
\end{parts}
%
\begin{solutionbox}{1.1in}
  A formation reaction is that of a compound from its standard state elements.
  \begin{enumerate}
    \item Two moles of the compound is formed, not one.
    \item It is not correctly balanced.
    \item Oxygen is not in its standard state.
  \end{enumerate}
\end{solutionbox}

\bigskip
\question A piece of $Ti$ with mass $m = 20.8$ g is heated in boiling water to $T_1 = 99.5^\circ$, then dropped into a calorimeter containing $V = 75$ mL of water at $T_2 = 21.7^\circ$. When equilibrium is reached, the temperature is $T_3 = 24.3^\circ$. Calculate the specific heat of titanium.
%
\begin{solutionbox}{1in}
  The heats are related via \begin{equation*}
    - m C_{Ti} \Delta T_{31} = V C \Delta T_{32} \implies C_{Ti} = 0.52
  \end{equation*}
  Therefore, the specific heat of titanium is 0.52 J/g $^\circ$C.
\end{solutionbox}

\bigskip
\question Using Hess' Law, calculate the heat of reaction for $A + B \to C$ given: \begin{equation*}
  2D \to B \quad \Delta H = 50 ,\qquad 2A + D \to C \quad \Delta H = -100 ,\qquad D \to A \quad \Delta H = 75
\end{equation*}
%
\begin{solutionbox}{2in}
  The net sum of the sub-reactions must be equal to the overall reaction. To achieve this, the first reaction is reversed such that $B \to 2D$ and $\Delta H = -50$. Adding this to the remaining two sub-reactions yields:
    \begin{align*}
      & 1. & B &\to 2D & \Delta H_1 &= -50 \\
      & 2. & 2A + D &\to C & \Delta H_2 &= -100 \\
      & 3. & D &\to A & \Delta H_3 &= 75 \\
      & \Sigma & A + B &\to C & \Delta H_t &= \Sigma \Delta H
    \end{align*}
    Therefore, the total enthalpy $\Delta H_{rxn} = \sum \Delta H = -75$ kJ / mol.
\end{solutionbox}

\bigskip
\question A metal has a cubic structure with density $\rho = 1.892$ g/cm$^3$, an atomic weight $M = 132.91$ g/mol, and a lattice parameter $a = 6.13 \cdot 10^{-8}$ cm. Determine the crystal structure of the metal.
%
\begin{solutionbox}{1.2in}
  The number of atoms per cell $n$ is given by \begin{equation*}
    \rho = \frac{nM}{a^3N_a} \implies n = \frac{\rho a^3 N_a}{M} = 2
  \end{equation*}
  Therefore, the structure is a body centered cubic as the number of atoms per cell is 2.
\end{solutionbox}

\bigskip
\question List and explain two factors that affect the conductivity of semiconductors.
%
\begin{solutionbox}{0.55in}
  Temperature and the number of impurities is proportional to conductivity in semiconductors, and opposite in conductors.
\end{solutionbox}

\bigskip
\question There are three types of polymer molecular structures.
\begin{parts}
  \part In which structures are the chains connected by non bonding interactions and covalent bonds?
  \part Which structure more easily forms a crystal? Explain.
  \part Which structure can be recycled?
  \part If a compound burns and does not melt, what type of structure is it?
\end{parts}
%
\begin{solutionbox}{1.5in}
  \begin{parts}
    \part Linear and branched polymers are connected by nonbonding interactions, whereas crosslinked polymers are connected by covalent bonds.
    \part Linear forms crystals more easily as there are no branches.
    \part Linear and branched polymers can be recycled as their interactions are broken when melted.
    \part It is either network or crosslinked.
  \end{parts}
\end{solutionbox}

\bigskip
\question Calculate the radius of a Pd atom in picometers, given $\rho = 12.02$ and such atom is an fcc unit cell.
%
\begin{solutionbox}{2.3in}
  The lattice parameter $a$ can be found via \[ a = \sqrt[3]{V} = \sqrt[3]{m/\rho} \quad\text{for}\quad m = nM/N_a \]
  Therefore, $a = 3.89 \cdot 10^{-8}$. Because the unit cell is fcc, \[r = a/2\sqrt{2} = 1.376 \cdot 10^{-8} \text{ cm} \]
  Converting to picometers from centimetres yields \[1.376 \cdot 10^{-8} \cdot 10^{8} \cdot 10^{2} = 138 \text{ pm} \]
  Therefore the radius is $r = 138$ pm.
\end{solutionbox}


\end{questions}

\end{document}
